%\documentclass{article}
\documentclass[a4paper,oneside,12pt]{amsart}
\usepackage[left=1.1cm,right=1.1cm,top=2cm,bottom=1cm,bindingoffset=0cm]{geometry}
%----------- fonts
\usepackage[english,russian]{babel}
\usepackage{fontspec}
\setmainfont{CMU Serif}
\usepackage{xeCJK}
\setCJKmainfont{TakaoMincho} % for \rmfamily
%------- another packages:
\usepackage{braket}
\usepackage{hyperref}
%------ the document begins:
\begin{document}
\setcounter{secnumdepth}{0}
\vspace{-9mm}
\title{The Sign of Four} \maketitle
\vspace{-2mm}

\begin{center}
\emph{by Arthur Conan Doyle}
\end{center}
\section{Chapter 4: The Story of the Bald-Headed Man}

{\large 

 {\bf Bald-Headed } --- \emph{ Плешивый }

{\bf sordid } [ˈsɔː.dɪd] ~ --- \emph{ противный }

{\bf ill-lit } --- \emph{ плохо освещенных }

{\bf worse } [wɜːs] ~ --- \emph{ хуже (от bad) }

{\bf furnished } [ˈfɜː.nɪʃt] ~ --- \emph{ снаряженный, оборудованный, экипированный }

{\bf threw } [θruː] ~ --- \emph{ бросал (не flung) }

{\bf blaze } [bleɪz] ~ --- \emph{ пламя }

{\bf glare } [ɡleər] ~ --- \emph{ сверкать }

{\bf bristle } [ˈbrɪs.əl] ~ --- \emph{ щетина }

{\bf fringe } [frɪndʒ] ~ --- \emph{ бахрома, край }

{\bf bald } [bɔːld] ~ --- \emph{ лысый }

{\bf scalp } [skælp] ~ --- \emph{ скальп }

{\bf shot } [ʃɒt] ~ --- \emph{ выстрел }

{\bf mountain-peak } --- \emph{ горная вершина }

{\bf fir-trees } --- \emph{ ели }

{\bf writhed } --- \emph{ корчился }

{\bf features } --- \emph{ Особенности }

{\bf perpetual } [pəˈpetʃ.u.əl] ~ --- \emph{ вечный }

{\bf jerk } [dʒɜːk] ~ --- \emph{ резкий толчок, подергивание; резко толкать, дергать }

{\bf scowling } --- \emph{ сердитый }

{\bf repose } [rɪˈpəʊz] ~ --- \emph{ отдыхать, лежать/отдых, покой }

{\bf given } [ˈɡɪv.ən] ~ --- \emph{ данный (от give) }

{\bf pendulous } [ˈpen.dʒə.ləs] ~ --- \emph{ подвешенный }

{\bf visible } [ˈvɪz.ə.bəl] ~ --- \emph{ видимый }

{\bf irregular } [ɪˈreɡ.jə.lər] ~ --- \emph{ неправильный }

{\bf strove } [strəʊv] ~ --- \emph{ норовил }

{\bf feebly } --- \emph{ бессильно }

{\bf conceal } [kənˈsiːl] ~ --- \emph{ скрывать }

{\bf constantly } [ˈkɒn.stənt.li] ~ --- \emph{ постоянно }

{\bf spite } [spaɪt] ~ --- \emph{ злоба }

{\bf obtrusive } [əbˈtruː.sɪv] ~ --- \emph{ навязчивость }

{\bf bald } [bɔːld] ~ --- \emph{ лысый }

{\bf impression } [ɪmˈpreʃ.ən] ~ --- \emph{ впечатление }

{\bf thirtieth } [ˈθɜː.ti.əθ] ~ --- \emph{ тридцатый }

} \vspace{6mm} {\Large 

We followed the Indian down a sordid and common passage, ill-lit and worse furnished, until he came to a door upon the right, which he threw open. A blaze of yellow light streamed out upon us, and in the centre of the glare there stood a small man with a very high head, a bristle of red hair all round the fringe of it, and a bald, shining scalp which shot out from among it like a mountain-peak from fir-trees. He writhed his hands together as he stood, and his features were in a perpetual jerk --- now smiling, now scowling, but never for an instant in repose. Nature had given him a pendulous lip, and a too visible line of yellow and irregular teeth, which he strove feebly to conceal by constantly passing his hand over the lower part of his face. In spite of his obtrusive baldness he gave the impression of youth. In point of fact, he had just turned his thirtieth year.
\\ } 

{\large 

 {\bf servant } [ˈsɜː.vənt] ~ --- \emph{ слуга }

{\bf gentlemen } --- \emph{ господа (от gentleman) }

{\bf pray } [preɪ] ~ --- \emph{ молиться }

{\bf sanctum } [ˈsæŋk.təm] ~ --- \emph{ святилище }

{\bf oasis } [əʊˈeɪ.sɪs] ~ --- \emph{ оазис }

{\bf howl } [haʊl] ~ --- \emph{ выть/вой }

{\bf astonished } [əˈstɒn.ɪʃt] ~ --- \emph{ удивленный }

{\bf apartment } [əˈpɑːt.mənt] ~ --- \emph{ квартира }

{\bf invited } --- \emph{ пригласил, приглашенный (от invite) }

{\bf sorry } [ˈsɒr.i] ~ --- \emph{ огорчённый, сожалеющий }

{\bf diamond } [ˈdaɪə.mənd] ~ --- \emph{ бриллиант }

{\bf brass } [brɑːs] ~ --- \emph{ латунь }

{\bf richest } --- \emph{ богатейший }

{\bf glossiest } --- \emph{ глянцевых }

{\bf curtains } --- \emph{ занавески }

{\bf tapestries } --- \emph{ гобелены }

{\bf draped } --- \emph{ задрапированный }

{\bf loop } [luːp] ~ --- \emph{ петля }

{\bf expose } [ɪkˈspəʊz] ~ --- \emph{ подвергать }

{\bf oriental } --- \emph{ восточный }

{\bf vase } [vɑːz] ~ --- \emph{ ваза }

{\bf carpet } [ˈkɑː.pɪt] ~ --- \emph{ ковёр }

{\bf amber } [ˈæm.bər] ~ --- \emph{ янтарь, янтарный }

{\bf sank } [sæŋk] ~ --- \emph{ тонул }

{\bf pleasantly } --- \emph{ приятно }

{\bf moss } [mɒs] ~ --- \emph{ мох }

{\bf tiger-skins } --- \emph{ тигровые шкуры }

{\bf thrown } --- \emph{ выброшены }

{\bf athwart } --- \emph{ через, вопреки (не through) }

{\bf eastern } [ˈiː.stən] ~ --- \emph{ восточный }

{\bf luxury } [ˈlʌk.ʃər.i] ~ --- \emph{ роскошь или наслаждение }

{\bf hookah } [ˈhʊk.ə] ~ --- \emph{ кальян }

{\bf lamp } [læmp] ~ --- \emph{ лампа }

{\bf fashion } [ˈfæʃ.ən] ~ --- \emph{ мода }

{\bf dove } [dʌv] ~ --- \emph{ голубь }

{\bf hung } [hʌŋ] ~ --- \emph{ вешал, был склонным }

{\bf invisible } [ɪnˈvɪz.ə.bəl] ~ --- \emph{ невидимый }

{\bf subtle } [ˈsʌt.əl] ~ --- \emph{ тонкий, неуловимый, искусный }

{\bf aromatic } [ˌær.əˈmæt.ɪk] ~ --- \emph{ ароматический }

{\bf odour } [ˈəʊ.dər] ~ --- \emph{ запах, аромат }

} \vspace{6mm} {\Large 

 
"Your servant, Miss Morstan," he kept repeating in a thin, high voice. "Your servant, gentlemen. Pray step into my little sanctum. A small place, miss, but furnished to my own liking. An oasis of art in the howling desert of South London."

We were all astonished by the appearance of the apartment into which he invited us. In that sorry house it looked as out of place as a diamond of the first water in a setting of brass. The richest and glossiest of curtains and tapestries draped the walls, looped back here and there to expose some richly mounted painting or Oriental vase. The carpet was of amber and black, so soft and so thick that the foot sank pleasantly into it, as into a bed of moss. Two great tiger-skins thrown athwart it increased the suggestion of Eastern luxury, as did a huge hookah which stood upon a mat in the corner. A lamp in the fashion of a silver dove was hung from an almost invisible golden wire in the centre of the room. As it burned it filled the air with a subtle and aromatic odour.
\\ } 

{\large 

 {\bf jerk } [dʒɜːk] ~ --- \emph{ резкий толчок, подергивание; резко толкать, дергать }

{\bf cried } [kraɪd] ~ --- \emph{ плакала }

{\bf excited } [ɪkˈsaɪ.tɪd] ~ --- \emph{ взволнованный, возбуждённый }

{\bf stethoscope } [ˈsteθ.ə.skəʊp] ~ --- \emph{ стетоскоп }

{\bf grave } [ɡreɪv] ~ --- \emph{ могила }

{\bf doubts } --- \emph{ Сомнения }

{\bf mitral } --- \emph{ митральный }

{\bf valve } [vælv] ~ --- \emph{ клапан, створка, радио электронная лампа }

{\bf aortic } --- \emph{ аортальный }

{\bf rely } --- \emph{ полагаться }

{\bf opinion } [əˈpɪn.jən] ~ --- \emph{ мнение }

} \vspace{6mm} {\Large 

 
"Mr. Thaddeus Sholto," said the little man, still jerking and smiling. "That is my name. You are Miss Morstan, of course. And these gentlemen ---"

"This is Mr. Sherlock Holmes, and this Dr. Watson."

"A doctor, eh?" cried he, much excited. "Have you your stethoscope? Might I ask you --- would you have the kindness? I have grave doubts as to my mitral valve, if you would be so very good. The aortic I may rely upon, but I should value your opinion upon the mitral."
\\ } 

{\large 

 {\bf unable } [ʌnˈeɪ.bəl] ~ --- \emph{ неспособный }

{\bf amiss } [əˈmɪs] ~ --- \emph{ неверно, не туда, некстати }

{\bf indeed } [ɪnˈdiːd] ~ --- \emph{ в самом деле }

{\bf ecstasy } [ˈek.stə.si] ~ --- \emph{ восторг, экстаз }

{\bf shiver } [ˈʃɪv.ər] ~ --- \emph{ дрожать/дрожь }

{\bf normal } [ˈnɔː.məl] ~ --- \emph{ нормальный }

{\bf uneasiness } [ʌnˈiː.zi.nəs] ~ --- \emph{ тревога, неловкость }

{\bf excuse } [ɪkˈskjuːz] ~ --- \emph{ извинение, оправдание/извинить }

{\bf anxiety } [æŋˈzaɪ.ə.ti] ~ --- \emph{ беспокойство, тревога }

{\bf remark } [rɪˈmɑːk] ~ --- \emph{ замечание/делать замечание }

{\bf airily } --- \emph{ легкомысленно }

{\bf suffer } [ˈsʌf.ər] ~ --- \emph{ страдать }

{\bf suspicions } --- \emph{ Подозрения }

{\bf delighted } [dɪˈlaɪ.tɪd] ~ --- \emph{ восхищался (от delight) }

{\bf unwarranted } [ʌnˈwɒr.ən.tɪd] ~ --- \emph{ необоснованный }

{\bf refrain } [rɪˈfreɪn] ~ --- \emph{ воздерживаться }

{\bf strain } [streɪn] ~ --- \emph{ напряжение, натяжение, испытание/напрягать, натягивать }

{\bf alive } [əˈlaɪv] ~ --- \emph{ живой, живущий, бодрый }

} \vspace{6mm} {\Large 

 
I listened to his heart, as requested, but was unable to find anything amiss, save, indeed, that he was in an ecstasy of fear, for he shivered from head to foot.

"It appears to be normal," I said. "You have no cause for uneasiness."

"You will excuse my anxiety, Miss Morstan," he remarked airily. "I am a great sufferer, and I have long had suspicions as to that valve. I am delighted to hear that they are unwarranted. Had your father, Miss Morstan, refrained from throwing a strain upon his heart, he might have been alive now."
\\ } 

{\large 

 {\bf struck } [strʌk] ~ --- \emph{ забастовка, удар (прошедшее время) }

{\bf callous } [ˈkæl.əs] ~ --- \emph{ черствый }

{\bf offhand } [ˌɒfˈhænd] ~ --- \emph{ экспромтом }

{\bf reference } [ˈref.ər.əns] ~ --- \emph{ ссылка, упоминание }

{\bf lip } [lɪp] ~ --- \emph{ губа }

{\bf whatever } [wɒtˈev.ər] ~ --- \emph{ что бы ни, какой бы ни, и т. д. }

{\bf escort } [ɪˈskɔːt] ~ --- \emph{ охрана (не guards) }

{\bf witness } [ˈwɪt.nəs] ~ --- \emph{ свидетель/быть очевидцем }

{\bf bold } [bəʊld] ~ --- \emph{ смелый, наглый, самоуверенный }

{\bf outsider } [ˌaʊtˈsaɪ.dər] ~ --- \emph{ посторонний }

{\bf police } [pəˈliːs] ~ --- \emph{ полиция }

{\bf everything } [ˈev.ri.θɪŋ] ~ --- \emph{ все }

{\bf satisfactorily } --- \emph{ удовлетворительно }

{\bf ourselves } [ˌaʊəˈselvz] ~ --- \emph{ себя, себе, сами }

{\bf interference } [ˌɪn.təˈfɪə.rəns] ~ --- \emph{ вмешательство }

{\bf annoy } [əˈnɔɪ] ~ --- \emph{ надоедать, докучать }

} \vspace{6mm} {\Large 

 
I could have struck the man across the face, so hot was I at this callous and offhand reference to so delicate a matter. Miss Morstan sat down, and her face grew white to the lips.

"I knew in my heart that he was dead," said she.

"I can give you every information," said he; "and, what is more, I can do you justice; and I will, too, whatever Brother Bartholomew may say. I am so glad to have your friends here not only as an escort to you but also as witnesses to what I am about to do and say. The three of us can show a bold front to Brother Bartholomew. But let us have no outsiders --- no police or officials. We can settle everything satisfactorily among ourselves without any interference. Nothing would annoy Brother Bartholomew more than any publicity."
\\ } 

{\large 

 {\bf blinked } --- \emph{ Мигал }

{\bf inquiringly } --- \emph{ вопросительно }

{\bf weak } [wiːk] ~ --- \emph{ слабый }

{\bf watery } [ˈwɔː.tər.i] ~ --- \emph{ водянистый, дождевой, бесцветный }

{\bf further } [ˈfɜː.ðər] ~ --- \emph{ дальшейший }

{\bf nod } [nɒd] ~ --- \emph{ кивать }

{\bf Tokay } --- \emph{ токайский }

{\bf flask } [flɑːsk] ~ --- \emph{ фляжка }

{\bf trust } [trʌst] ~ --- \emph{ доверять }

{\bf tobacco-smoke } --- \emph{ табачный дым }

{\bf balsamic } --- \emph{ бальзамического }

{\bf odour } [ˈəʊ.dər] ~ --- \emph{ запах, аромат }

{\bf eastern } [ˈiː.stən] ~ --- \emph{ восточный }

{\bf tobacco } [təˈbæk.əʊ] ~ --- \emph{ табак }

{\bf nervous } [ˈnɜː.vəs] ~ --- \emph{ нервный }

{\bf hookah } [ˈhʊk.ə] ~ --- \emph{ кальян }

{\bf invaluable } [ɪnˈvæl.jə.bəl] ~ --- \emph{ бесценный }

{\bf sedative } [ˈsed.ə.tɪv] ~ --- \emph{ успокоительное }

} \vspace{6mm} {\Large 

 
He sat down upon a low settee and blinked at us inquiringly with his weak, watery blue eyes.

"For my part," said Holmes, "whatever you may choose to say will go no further."

I nodded to show my agreement.

"That is well! That is well!" said he. "May I offer you a glass of Chianti, Miss Morstan? Or of Tokay? I keep no other wines. Shall I open a flask? No? Well, then, I trust that you have no objection to tobacco-smoke, to the balsamic odour of the Eastern tobacco. I am a little nervous, and I find my hookah an invaluable sedative."
\\ } 

{\large 

 {\bf applied } [əˈplaɪd] ~ --- \emph{ прикладной }

{\bf tap } [tæp] ~ --- \emph{ стучать, похлопывать/стук }

{\bf bowl } [bəʊl] ~ --- \emph{ миска, чаша }

{\bf smoke } [sməʊk] ~ --- \emph{ курить }

{\bf bubbled } --- \emph{ ыми }

{\bf merry } [ˈmer.i] ~ --- \emph{ весёлый, радостный }

{\bf rose-water } [ˈrəʊz ˌwɔː.tər] ~ --- \emph{ розово-вода }

{\bf semicircle } [ˈsem.iˌsɜː.kəl] ~ --- \emph{ полукруг }

{\bf advanced } [ədˈvɑːnst] ~ --- \emph{ передовой }

{\bf chin } [tʃɪn] ~ --- \emph{ подбородок }

{\bf jerky } [ˈdʒɜː.ki] ~ --- \emph{ вяленое мясо }

{\bf fellow } [ˈfel.əʊ] ~ --- \emph{ товарищ или просто чувак }

{\bf puff } [pʌf] ~ --- \emph{ дымок, дуновение (ветра); дымить, попыхивать }

{\bf uneasy } [ʌnˈiː.zi] ~ --- \emph{ неспокойный }

} \vspace{6mm} {\Large 

 
He applied a taper to the great bowl, and the smoke bubbled merrily through the rose-water. We sat all three in a semicircle, with our heads advanced and our chins upon our hands, while the strange, jerky little fellow, with his high, shining head, puffed uneasily in the centre.
\\ } 

{\large 

 {\bf determined } [dɪˈtɜː.mɪnd] ~ --- \emph{ определённый или решительный }

{\bf communication } [kəˌmjuː.nɪˈkeɪ.ʃən] ~ --- \emph{ общение }

{\bf given } [ˈɡɪv.ən] ~ --- \emph{ данный (от give) }

{\bf address } [əˈdres] ~ --- \emph{ обращаться/обращение, выступление }

{\bf disregard } [ˌdɪs.rɪˈɡɑːd] ~ --- \emph{ не обращать внимания/равнодушие, игнорирование }

{\bf unpleasant } [ʌnˈplez.ənt] ~ --- \emph{ неприятный }

{\bf liberty } [ˈlɪb.ə.ti] ~ --- \emph{ свобода }

{\bf therefore } [ˈðeə.fɔːr] ~ --- \emph{ поэтому, следовательно }

{\bf appointment } [əˈpɔɪnt.mənt] ~ --- \emph{ назначение, определение }

{\bf william } --- \emph{ Уильям }

{\bf confidence } [ˈkɒn.fɪ.dəns] ~ --- \emph{ доверие, уверенность }

{\bf discretion } [dɪˈskreʃ.ən] ~ --- \emph{ благоразумие }

{\bf dissatisfy } --- \emph{ не удовлетворять }

{\bf excuse } [ɪkˈskjuːz] ~ --- \emph{ извинение, оправдание/извинить }

{\bf precaution } [prɪˈkɔː.ʃən] ~ --- \emph{ предосторожность или предостережение }

{\bf somewhat } [ˈsʌm.wɒt] ~ --- \emph{ отчасти, до некоторой степени }

{\bf retiring } [rɪˈtaɪə.rɪŋ] ~ --- \emph{ скромный }

{\bf refined } [rɪˈfaɪnd] ~ --- \emph{ утонченный }

{\bf tastes } --- \emph{ имеет вкус (от taste) }

{\bf unaesthetic } --- \emph{ неэстетический }

{\bf policeman } [pəˈliːs.mən] ~ --- \emph{ полицейский }

{\bf shrink } [ʃrɪŋk] ~ --- \emph{ сокращаться, сжиматься, усыхать }

{\bf rough } [rʌf] ~ --- \emph{ грубый }

{\bf mat } [mæt] ~ --- \emph{ коврик (не rug) }

{\bf seldom } [ˈsel.dəm] ~ --- \emph{ редко }

{\bf contact } [ˈkɒn.tækt] ~ --- \emph{ отношение, связь, взаимодействие/соприкосаться, взаимодействовать }

{\bf atmosphere } [ˈæt.mə.sfɪər] ~ --- \emph{ атмосфера }

{\bf elegance } --- \emph{ элегантность }

{\bf myself } [maɪˈself] ~ --- \emph{ себя, меня самого, себе, сам }

{\bf patron } [ˈpeɪ.trən] ~ --- \emph{ покровитель }

{\bf weakness } [ˈwiːk.nəs] ~ --- \emph{ слабость }

{\bf landscape } [ˈlænd.skeɪp] ~ --- \emph{ пейзаж, ладшафт }

{\bf genuine } [ˈdʒen.ju.ɪn] ~ --- \emph{ подлинный, истинный }

{\bf Corot } --- \emph{ Коро }

{\bf connoisseur } [ˌkɒn.əˈsɜːr] ~ --- \emph{ знаток }

{\bf doubt } [daʊt] ~ --- \emph{ сомнение/сомневаться }

{\bf Rosa } --- \emph{ Роза }

{\bf cannot } [ˈkæn.ɒt] ~ --- \emph{ не уметь (отрицательная форма глагола can) }

{\bf Bouguereau } --- \emph{ Бугро }

{\bf French } --- \emph{ французский }

} \vspace{6mm} {\Large 

 
"When I first determined to make this communication to you," said he, "I might have given you my address; but I feared that you might disregard my request and bring unpleasant people with you. I took the liberty, therefore, of making an appointment in such a way that my man Williams might be able to see you first. I have complete confidence in his discretion, and he had orders, if he were dissatisfied, to proceed no further in the matter. You will excuse these precautions, but I am a man of somewhat retiring, and I might even say refined, tastes, and there is nothing more unaesthetic than a policeman. I have a natural shrinking from all forms of rough materialism. I seldom come in contact with the rough crowd. I live, as you see, with some little atmosphere of elegance around me. I may call myself a patron of the arts. It is my weakness. The landscape is a genuine Corot, and though a connoisseur might perhaps throw a doubt upon that Salvator Rosa, there cannot be the least question about the Bouguereau. I am partial to the modern French school."
\\ } 

{\large 

 {\bf desire } [dɪˈzaɪər] ~ --- \emph{ желать/желание }

{\bf Norwood } --- \emph{ Норвуд }

{\bf angry } [ˈæŋ.ɡri] ~ --- \emph{ сердитый, раздражённый }

{\bf terrible } [ˈter.ə.bəl] ~ --- \emph{ ужасный }

{\bf Norwood } --- \emph{ Норвуд }

{\bf vent } [vent] ~ --- \emph{ выпускать, испускать }

{\bf remark } [rɪˈmɑːk] ~ --- \emph{ замечание/делать замечание }

{\bf cried } [kraɪd] ~ --- \emph{ плакала }

{\bf ignorant } [ˈɪɡ.nər.ənt] ~ --- \emph{ невежественный }

{\bf John } --- \emph{ Джон }

{\bf army } [ˈɑː.mi] ~ --- \emph{ армия }

{\bf retired } [rɪˈtaɪəd] ~ --- \emph{ отставной }

{\bf eleven } [ɪˈlev.ən] ~ --- \emph{ одиннадцать }

{\bf Norwood } --- \emph{ Норвуд }

{\bf prosper } [ˈprɒs.pər] ~ --- \emph{ процветать }

{\bf valuable } [ˈvæl.jə.bəl] ~ --- \emph{ ценный, дорогой }

{\bf curiosity } [ˌkjʊə.riˈɒs.ə.ti] ~ --- \emph{ любопытство }

{\bf native } [ˈneɪ.tɪv] ~ --- \emph{ родной }

{\bf servant } [ˈsɜː.vənt] ~ --- \emph{ слуга }

{\bf luxury } [ˈlʌk.ʃər.i] ~ --- \emph{ роскошь или наслаждение }

{\bf twin-brother } --- \emph{ Близнецы }

} \vspace{6mm} {\Large 

 
"You will excuse me, Mr. Sholto," said Miss Morstan, "but I am here at your request to learn something which you desire to tell me. It is very late, and I should desire the interview to be as short as possible."

"At the best it must take some time," he answered; "for we shall certainly have to go to Norwood and see Brother Bartholomew. We shall all go and try if we can get the better of Brother Bartholomew. He is very angry with me for taking the course which has seemed right to me. I had quite high words with him last night. You cannot imagine what a terrible fellow he is when he is angry."

"If we are to go to Norwood, it would perhaps be as well to start at once," I ventured to remark.

He laughed until his ears were quite red.

"That would hardly do," he cried. "I don't know what he would say if I brought you in that sudden way. No, I must prepare you by showing you how we all stand to each other. In the first place, I must tell you that there are several points in the story of which I am myself ignorant. I can only lay the facts before you as far as I know them myself.

"My father was, as you may have guessed, Major John Sholto, once of the Indian Army. He retired some eleven years ago and came to live at Pondicherry Lodge in Upper Norwood. He had prospered in India and brought back with him a considerable sum of money, a large collection of valuable curiosities, and a staff of native servants. With these advantages he bought himself a house, and lived in great luxury. My twin-brother Bartholomew and I were the only children.
\\ } 

{\large 

 {\bf sensation } [senˈseɪ.ʃən] ~ --- \emph{ чувство }

{\bf caused } --- \emph{ вызванный }

{\bf disappearance } [ˌdɪs.əˈpɪə.rəns] ~ --- \emph{ исчезновение }

{\bf detail } [ˈdiː.teɪl] ~ --- \emph{ подробность, деталь }

{\bf freely } [ˈfriː.li] ~ --- \emph{ свободно }

{\bf presence } [ˈprez.əns] ~ --- \emph{ присутствие }

{\bf suspect } [səˈspekt] ~ --- \emph{ подозревать/подозреваемый }

{\bf secret } [ˈsiː.krət] ~ --- \emph{ тайна }

{\bf hidden } [ˈhɪd.ən] ~ --- \emph{ скрытый }

{\bf breast } [brest] ~ --- \emph{ грудь }

{\bf alone } [əˈləʊn] ~ --- \emph{ один, одинокий, сам }

} \vspace{6mm} {\Large 

 
"I very well remember the sensation which was caused by the disappearance of Captain Morstan. We read the details in the papers, and knowing that he had been a friend of our father's we discussed the case freely in his presence. He used to join in our speculations as to what could have happened. Never for an instant did we suspect that he had the whole secret hidden in his own breast, that of all men he alone knew the fate of Arthur Morstan.
\\ } 

{\large 

 {\bf mystery } [ˈmɪs.tər.i] ~ --- \emph{ тайна }

{\bf positive } [ˈpɒz.ə.tɪv] ~ --- \emph{ положительный }

{\bf overhung } [ˌəʊ.vəˈhʌŋ] ~ --- \emph{ нависающий }

{\bf employed } [ɪmˈplɔɪd] ~ --- \emph{ Используемый }

{\bf prize-fighters } --- \emph{ приз Пожарным }

{\bf william } --- \emph{ Уильям }

{\bf drove } [drəʊv] ~ --- \emph{ водил машину, ехал на машине (от drive) }

{\bf tonight } [təˈnaɪt] ~ --- \emph{ сегодня вечером }

{\bf lightweight } [ˈlaɪt.weɪt] ~ --- \emph{ легкий }

{\bf champion } [ˈtʃæm.pi.ən] ~ --- \emph{ чемпион }

{\bf aversion } [əˈvɜː.ʃən] ~ --- \emph{ отвращение }

{\bf occasion } [əˈkeɪ.ʒən] ~ --- \emph{ случай, возможность }

{\bf revolver } [rɪˈvɒl.vər] ~ --- \emph{ револьвер }

{\bf wooden-legged } --- \emph{ деревянные ноги }

{\bf harmless } [ˈhɑːm.ləs] ~ --- \emph{ безвредный }

{\bf tradesman } [ˈtreɪdz.mən] ~ --- \emph{ ремесленник }

{\bf canvassing } --- \emph{ понуждение }

{\bf hush } [hʌʃ] ~ --- \emph{ водворять тишину }

{\bf mere } [mɪər] ~ --- \emph{ простой, явный }

{\bf whim } [wɪm] ~ --- \emph{ прихоть, причуда }

{\bf opinion } [əˈpɪn.jən] ~ --- \emph{ мнение }

} \vspace{6mm} {\Large 

 
"We did know, however, that some mystery, some positive danger, overhung our father. He was very fearful of going out alone, and he always employed two prize-fighters to act as porters at Pondicherry Lodge. Williams, who drove you tonight, was one of them. He was once lightweight champion of England. Our father would never tell us what it was he feared, but he had a most marked aversion to men with wooden legs. On one occasion he actually fired his revolver at a wooden-legged man, who proved to be a harmless tradesman canvassing for orders. We had to pay a large sum to hush the matter up. My brother and I used to think this a mere whim of my father's, but events have since led us to change our opinion.
\\ } 

{\large 

 {\bf shock } [ʃɒk] ~ --- \emph{ потрясать/потрясение, шок }

{\bf faint } [feɪnt] ~ --- \emph{ слабый, недостаточный }

{\bf breakfast-table } --- \emph{ завтрак стол }

{\bf sicken } [ˈsɪk.ən] ~ --- \emph{ испытывать отвращение или заболевать }

{\bf discover } [dɪˈskʌv.ər] ~ --- \emph{ открывать, узнавать }

{\bf scrawling } --- \emph{ царапать }

{\bf suffer } [ˈsʌf.ər] ~ --- \emph{ страдать }

{\bf enlarged } --- \emph{ увеличенный }

{\bf spleen } [spliːn] ~ --- \emph{ селезенка }

{\bf rapidly } --- \emph{ быстро }

{\bf worse } [wɜːs] ~ --- \emph{ хуже (от bad) }

{\bf April } --- \emph{ апрель }

{\bf informed } [ɪnˈfɔːmd] ~ --- \emph{ осведомленный }

{\bf communication } [kəˌmjuː.nɪˈkeɪ.ʃən] ~ --- \emph{ общение }

} \vspace{6mm} {\Large 

 
"Early in 1882 my father received a letter from India which was a great shock to him. He nearly fainted at the breakfast-table when he opened it, and from that day he sickened to his death. What was in the letter we could never discover, but I could see as he held it that it was short and written in a scrawling hand. He had suffered for years from an enlarged spleen, but he now became rapidly worse, and towards the end of April we were informed that he was beyond all hope, and that he wished to make a last communication to us.
\\ } 

{\large 

 {\bf prop } [prɒp] ~ --- \emph{ подпорка/подпирать }

{\bf pillow } [ˈpɪl.əʊ] ~ --- \emph{ подушка }

{\bf breathing } [ˈbriː.ðɪŋ] ~ --- \emph{ дыхание }

{\bf besought } [bɪˈsɔːt] ~ --- \emph{ умолять }

{\bf lock } [lɒk] ~ --- \emph{ замок/запирать }

{\bf grasp } [ɡrɑːsp] ~ --- \emph{ хватать/хватка }

{\bf remarkable } [rɪˈmɑː.kə.bəl] ~ --- \emph{ замечательный, выдающийся }

{\bf broken } [ˈbrəʊ.kən] ~ --- \emph{ разбитый, сломанный, нарушенный, ломаный (о языке) }

{\bf emotion } [ɪˈməʊ.ʃən] ~ --- \emph{ возбуждение }

{\bf pain } [peɪn] ~ --- \emph{ боль }

} \vspace{6mm} {\Large 

 
"When we entered his room he was propped up with pillows and breathing heavily. He besought us to lock the door and to come upon either side of the bed. Then grasping our hands he made a remarkable statement to us in a voice which was broken as much by emotion as by pain. I shall try and give it to you in his own very words.
\\ } 

{\large 

 {\bf weigh } [weɪ] ~ --- \emph{ весить }

{\bf supreme } [suːˈpriːm] ~ --- \emph{ верховный, высший }

{\bf orphan } [ˈɔː.fən] ~ --- \emph{ сирота }

{\bf cursed } [kɜːst] ~ --- \emph{ проклят }

{\bf greed } [ɡriːd] ~ --- \emph{ жадность }

{\bf beset } [bɪˈset] ~ --- \emph{ окружать }

{\bf withheld } --- \emph{ удержан }

{\bf treasure } [ˈtreʒ.ər] ~ --- \emph{ сокровище }

{\bf myself } [maɪˈself] ~ --- \emph{ себя, меня самого, себе, сам }

{\bf blind } [blaɪnd] ~ --- \emph{ слепой }

{\bf foolish } [ˈfuː.lɪʃ] ~ --- \emph{ глупый }

{\bf avarice } [ˈæv.ər.ɪs] ~ --- \emph{ алчность }

{\bf possession } [pəˈzeʃ.ən] ~ --- \emph{ владение, собственность }

{\bf chaplet } --- \emph{ четки }

{\bf tip } [tɪp] ~ --- \emph{ кончик }

{\bf pearls } --- \emph{ Жемчуг }

{\bf beside } [bɪˈsaɪd] ~ --- \emph{ рядом }

{\bf quinine-bottle } --- \emph{ хинин-бутылка }

{\bf Agra } --- \emph{ Агра }

{\bf recovered } --- \emph{ выздоровел }

} \vspace{6mm} {\Large 

 
" 'I have only one thing,' he said, 'which weighs upon my mind at this supreme moment. It is my treatment of poor Morstan's orphan. The cursed greed which has been my besetting sin through life has withheld from her the treasure, half at least of which should have been hers. And yet I have made no use of it myself, so blind and foolish a thing is avarice. The mere feeling of possession has been so dear to me that I could not bear to share it with another. See that chaplet tipped with pearls beside the quinine-bottle. Even that I could not bear to part with, although I had got it out with the design of sending it to her. You, my sons, will give her a fair share of the Agra treasure. But send her nothing --- not even the chaplet --- until I am gone. After all, men have been as bad as this and have recovered.
\\ } 

{\large 

 {\bf continued } [kənˈtɪn.juːd] ~ --- \emph{ продолжал (от continue) }

{\bf weak } [wiːk] ~ --- \emph{ слабый }

{\bf concealed } --- \emph{ Скрытый }

{\bf alone } [əˈləʊn] ~ --- \emph{ один, одинокий, сам }

{\bf circumstances } --- \emph{ обстоятельства, условия }

{\bf arrival } [əˈraɪ.vəl] ~ --- \emph{ прибытие }

{\bf admitted } --- \emph{ Признанный(допущенный) }

{\bf faithful } [ˈfeɪθ.fəl] ~ --- \emph{ верный, преданный }

{\bf Chowdar } --- \emph{ густая похлебка }

{\bf opinion } [əˈpɪn.jən] ~ --- \emph{ мнение }

{\bf sprung } [sprʌŋ] ~ --- \emph{ Прыгал (от слова spring) }

{\bf paroxysm } [ˈpær.ɒk.sɪ.zəm] ~ --- \emph{ пароксизм }

{\bf dusky } [ˈdʌs.ki] ~ --- \emph{ сумеречный }

{\bf backward } [ˈbæk.wəd] ~ --- \emph{ обратный/назад, наоборот }

{\bf treasure-chest } --- \emph{ сундук с сокровищами }

{\bf stooped } --- \emph{ Наклоненный }

{\bf horror } [ˈhɒr.ər] ~ --- \emph{ ужас или отвращение }

} \vspace{6mm} {\Large 

 
" 'I will tell you how Morstan died,' he continued. 'He had suffered for years from a weak heart, but he concealed it from every one. I alone knew it. When in India, he and I, through a remarkable chain of circumstances, came into possession of a considerable treasure. I brought it over to England, and on the night of Morstan's arrival he came straight over here to claim his share. He walked over from the station and was admitted by my faithful old Lal Chowdar, who is now dead. Morstan and I had a difference of opinion as to the division of the treasure, and we came to heated words. Morstan had sprung out of his chair in a paroxysm of anger, when he suddenly pressed his hand to his side, his face turned a dusky hue, and he fell backward, cutting his head against the corner of the treasure-chest. When I stooped over him I found, to my horror, that he was dead.
\\ } 

{\large 

 {\bf distracted } [dɪˈstræk.tɪd] ~ --- \emph{ Отвлеченный }

{\bf impulse } [ˈɪm.pʌls] ~ --- \emph{ толчок, побуждение }

{\bf assistance } [əˈsɪs.təns] ~ --- \emph{ помощь, содействие }

{\bf recognize } [ˈrek.əɡ.naɪz] ~ --- \emph{ признавать, распознавать }

{\bf accused } [əˈkjuːzd] ~ --- \emph{ Обвиняемый }

{\bf murder } [ˈmɜː.dər] ~ --- \emph{ убийство/убивать }

{\bf quarrel } [ˈkwɒr.əl] ~ --- \emph{ ссора/ссориться }

{\bf gash } [ɡæʃ] ~ --- \emph{ запил }

{\bf inquiry } [ɪnˈkwaɪə.ri] ~ --- \emph{ вопрос, запрос или исследование }

{\bf anxious } [ˈæŋk.ʃəs] ~ --- \emph{ озабоченный, беспокоящийся }

{\bf secret } [ˈsiː.krət] ~ --- \emph{ тайна }

{\bf soul } [səʊl] ~ --- \emph{ душа }

{\bf necessity } [nəˈses.ə.ti] ~ --- \emph{ необходимость }

} \vspace{6mm} {\Large 

 
" 'For a long time I sat half distracted, wondering what I should do. My first impulse was, of course, to call for assistance; but I could not but recognize that there was every chance that I would be accused of his murder. His death at the moment of a quarrel, and the gash in his head, would be black against me. Again, an official inquiry could not be made without bringing out some facts about the treasure, which I was particularly anxious to keep secret. He had told me that no soul upon earth knew where he had gone. There seemed to be no necessity why any soul ever should know.
\\ } 

{\large 

 {\bf ponder } [ˈpɒn.dər] ~ --- \emph{ обдумывать, взвешивать }

{\bf servant } [ˈsɜː.vənt] ~ --- \emph{ слуга }

{\bf Chowdar } --- \emph{ густая похлебка }

{\bf doorway } [ˈdɔː.weɪ] ~ --- \emph{ дверной проём или путь к чему-либо (перен.) }

{\bf stole } [stəʊl] ~ --- \emph{ украл }

{\bf bolt } [bəʊlt] ~ --- \emph{ молния }

{\bf sahib } --- \emph{ сагиб }

{\bf hide } [haɪd] ~ --- \emph{ скрывать }

{\bf wiser } --- \emph{ мудрее }

{\bf Chowdar } --- \emph{ густая похлебка }

{\bf shook } [ʃʊk] ~ --- \emph{ трёс }

{\bf lip } [lɪp] ~ --- \emph{ губа }

{\bf seal } [siːl] ~ --- \emph{ печать/опечатывать }

{\bf asleep } [əˈsliːp] ~ --- \emph{ спящий }

{\bf innocence } [ˈɪn.ə.səns] ~ --- \emph{ невиновность }

{\bf twelve } [twelv] ~ --- \emph{ двенадцать }

{\bf tradesman } [ˈtreɪdz.mən] ~ --- \emph{ ремесленник }

{\bf jury-box } [ˈdʒʊə.ri ˌbɒks] ~ --- \emph{ коробка жюри }

{\bf Chowdar } --- \emph{ густая похлебка }

{\bf disposed } [dɪˈspəʊzd] ~ --- \emph{ склонный, расположенный, имеющийся под рукой (не inclined) }

{\bf mysterious } [mɪˈstɪə.ri.əs] ~ --- \emph{ таинственный, непостижимый }

{\bf disappearance } [ˌdɪs.əˈpɪə.rəns] ~ --- \emph{ исчезновение }

{\bf blame } [bleɪm] ~ --- \emph{ порицать, винить/упрёк }

{\bf fault } [fɒlt] ~ --- \emph{ вина }

{\bf clung } [klʌŋ] ~ --- \emph{ Цеплялся }

{\bf therefore } [ˈðeə.fɔːr] ~ --- \emph{ поэтому, следовательно }

{\bf restitution } [ˌres.tɪˈtʃuː.ʃən] ~ --- \emph{ реституция }

{\bf hidden } [ˈhɪd.ən] ~ --- \emph{ скрытый }

} \vspace{6mm} {\Large 

 
" 'I was still pondering over the matter, when, looking up, I saw my servant, Lal Chowdar, in the doorway. He stole in and bolted the door behind him. "Do not fear, sahib," he said; "no one need know that you have killed him. Let us hide him away, and who is the wiser?" "I did not kill him," said I. Lal Chowdar shook his head and smiled. "I heard it all, sahib," said he; "I heard you quarrel, and I heard the blow. But my lips are sealed. All are asleep in the house. Let us put him away together." That was enough to decide me. If my own servant could not believe my innocence, how could I hope to make it good before twelve foolish tradesmen in a jury-box? Lal Chowdar and I disposed of the body that night, and within a few days the London papers were full of the mysterious disappearance of Captain Morstan. You will see from what I say that I can hardly be blamed in the matter. My fault lies in the fact that we concealed not only the body but also the treasure and that I have clung to Morstan's share as well as to my own. I wish you, therefore, to make restitution. Put your ears down to my mouth. The treasure is hidden in ---'
\\ } 

{\large 

 {\bf horrible } [ˈhɒr.ə.bəl] ~ --- \emph{ страшный, отвратительный }

{\bf expression } [ɪkˈspreʃ.ən] ~ --- \emph{ выражение }

{\bf yell } [jel] ~ --- \emph{ пронзительный крик/кричать }

{\bf forget } [fəˈɡet] ~ --- \emph{ забыть }

{\bf Christ's } --- \emph{ Христовы }

{\bf sake } [seɪk] ~ --- \emph{ ради, для }

{\bf gaze } [ɡeɪz] ~ --- \emph{ пристальный взгляд/пристально глядеть }

{\bf fixed } [fɪkst] ~ --- \emph{ назначил (от fix) }

{\bf whitening } [ˈwaɪ.tən.ɪŋ] ~ --- \emph{ беление }

{\bf beard } [bɪəd] ~ --- \emph{ борода }

{\bf hairy } [ˈheə.ri] ~ --- \emph{ волосатый }

{\bf cruel } [ˈkruː.əl] ~ --- \emph{ жестокий }

{\bf concentrated } [ˈkɒn.sən.treɪ.tɪd] ~ --- \emph{ концентрированный }

{\bf malevolence } --- \emph{ недоброжелательность }

{\bf rush } [rʌʃ] ~ --- \emph{ торопиться, мчаться }

{\bf pulse } [pʌls] ~ --- \emph{ пульс, пульсировать }

{\bf ceased } --- \emph{ Прекратил }

} \vspace{6mm} {\Large 

 
"At this instant a horrible change came over his expression; his eyes stared wildly, his jaw dropped, and he yelled in a voice which I can never forget, 'Keep him out! For Christ's sake keep him out!' We both stared round at the window behind us upon which his gaze was fixed. A face was looking in at us out of the darkness. We could see the whitening of the nose where it was pressed against the glass. It was a bearded, hairy face, with wild cruel eyes and an expression of concentrated malevolence. My brother and I rushed towards the window, but the man was gone. When we returned to my father his head had dropped and his pulse had ceased to beat.
\\ } 

{\large 

 {\bf intruder } [ɪnˈtruː.dər] ~ --- \emph{ нарушитель }

{\bf footmark } --- \emph{ след }

{\bf visible } [ˈvɪz.ə.bəl] ~ --- \emph{ видимый }

{\bf flower-bed } [ˈflaʊ.ə bed] ~ --- \emph{ Кумба }

{\bf trace } [treɪs] ~ --- \emph{ след/следить }

{\bf imagination } [ɪˌmædʒ.ɪˈneɪ.ʃən] ~ --- \emph{ воображение }

{\bf conjured } --- \emph{ колдовал }

{\bf fierce } [fɪəs] ~ --- \emph{ свирепый }

{\bf proof } [pruːf] ~ --- \emph{ доказательство }

{\bf cupboard } [ˈkʌb.əd] ~ --- \emph{ буфет }

{\bf rifled } --- \emph{ нарезной }

{\bf chest } [tʃest] ~ --- \emph{ сундук, ящик }

{\bf torn } [tɔːn] ~ --- \emph{ рваный }

{\bf scrawled } --- \emph{ нацарапал }

{\bf judge } [dʒʌdʒ] ~ --- \emph{ составлять мнение, судить }

{\bf none } [nʌn] ~ --- \emph{ никто, никакой, ни один }

{\bf stolen } [ˈstəʊ.lən] ~ --- \emph{ украденный (от steal) }

{\bf everything } [ˈev.ri.θɪŋ] ~ --- \emph{ все }

{\bf naturally } [ˈnætʃ.ər.əl.i] ~ --- \emph{ естественно }

{\bf associated } [əˈsəʊ.si.eɪ.tɪd] ~ --- \emph{ связанный }

{\bf peculiar } [pɪˈkjuː.li.ər] ~ --- \emph{ определённый, специфический }

{\bf incident } [ˈɪn.sɪ.dənt] ~ --- \emph{ случай, происществие, эпизод }

{\bf haunted } [ˈhɔːn.tɪd] ~ --- \emph{ Часто посещаемый }

{\bf mystery } [ˈmɪs.tər.i] ~ --- \emph{ тайна }

} \vspace{6mm} {\Large 

 
"We searched the garden that night but found no sign of the intruder save that just under the window a single footmark was visible in the flower-bed. But for that one trace, we might have thought that our imaginations had conjured up that wild, fierce face. We soon, however, had another and a more striking proof that there were secret agencies at work all round us. The window of my father's room was found open in the morning, his cupboards and boxes had been rifled, and upon his chest was fixed a torn piece of paper with the words 'The sign of the four' scrawled across it. What the phrase meant or who our secret visitor may have been, we never knew. As far as we can judge, none of my father's property had been actually stolen, though everything had been turned out. My brother and I naturally associated this peculiar incident with the fear which haunted my father during his life, but it is still a complete mystery to us."
\\ } 

{\large 

 {\bf relight } --- \emph{ снова загореться }

{\bf hookah } [ˈhʊk.ə] ~ --- \emph{ кальян }

{\bf puff } [pʌf] ~ --- \emph{ дымок, дуновение (ветра); дымить, попыхивать }

{\bf absorbed } [əbˈzɔːbd] ~ --- \emph{ увлечённый чем-либо }

{\bf extraordinary } [ɪkˈstrɔː.dɪn.ər.i] ~ --- \emph{ необычайный, выдающийся }

{\bf narrative } [ˈnær.ə.tɪv] ~ --- \emph{ повествование, рассказ }

{\bf account } [əˈkaʊnt] ~ --- \emph{ счёт в банке }

{\bf faint } [feɪnt] ~ --- \emph{ слабый, недостаточный }

{\bf rallied } --- \emph{ сплотились }

{\bf poured } --- \emph{ Лил }

{\bf Venetian } --- \emph{ венецианец }

{\bf carafe } [kəˈræf] ~ --- \emph{ графин }

{\bf side-table } --- \emph{ столик }

{\bf lean } [liːn] ~ --- \emph{ худой }

{\bf abstract } [ˈæb.strækt] ~ --- \emph{ отвлеченный, абстрактный }

{\bf lid } [lɪd] ~ --- \emph{ крышка }

{\bf drawn } [drɔːn] ~ --- \emph{ нарисованный (от draw) }

{\bf glittering } [ˈɡlɪt.ər.ɪŋ] ~ --- \emph{ Сверкание }

{\bf glanced } --- \emph{ Поглядел }

{\bf complain } --- \emph{ жаловаться }

{\bf commonplaceness } --- \emph{ обыденности }

{\bf sagacity } --- \emph{ проницательность }

{\bf utmost } [ˈʌt.məʊst] ~ --- \emph{ крайний, предельный, величайший }

{\bf obvious } [ˈɒb.vi.əs] ~ --- \emph{ очевидный }

{\bf pride } [praɪd] ~ --- \emph{ гордость }

{\bf produced } --- \emph{ производил (от produce) }

{\bf continued } [kənˈtɪn.juːd] ~ --- \emph{ продолжал (от continue) }

{\bf puff } [pʌf] ~ --- \emph{ дымок, дуновение (ветра); дымить, попыхивать }

{\bf overgrown } [ˌəʊ.vəˈɡrəʊn] ~ --- \emph{ заросший }

{\bf pipe } [paɪp] ~ --- \emph{ свисток }

} \vspace{6mm} {\Large 

 
The little man stopped to relight his hookah and puffed thoughtfully for a few moments. We had all sat absorbed, listening to his extraordinary narrative. At the short account of her father's death Miss Morstan had turned deadly white, and for a moment I feared that she was about to faint. She rallied, however, on drinking a glass of water which I quietly poured out for her from a Venetian carafe upon the side-table. Sherlock Holmes leaned back in his chair with an abstracted expression and the lids drawn low over his glittering eyes. As I glanced at him I could not but think how on that very day he had complained bitterly of the commonplaceness of life. Here at least was a problem which would tax his sagacity to the utmost. Mr. Thaddeus Sholto looked from one to the other of us with an obvious pride at the effect which his story had produced and then continued between the puffs of his overgrown pipe.
\\ } 

{\large 

 {\bf excited } [ɪkˈsaɪ.tɪd] ~ --- \emph{ взволнованный, возбуждённый }

{\bf treasure } [ˈtreʒ.ər] ~ --- \emph{ сокровище }

{\bf spoken } [ˈspəʊ.kən] ~ --- \emph{ сказанный }

{\bf delved } --- \emph{ углубился }

{\bf discovering } --- \emph{ обнаружение }

{\bf whereabouts } [ˈweə.rə.baʊts] ~ --- \emph{ местонахождение }

{\bf madden } [ˈmæd.ən] ~ --- \emph{ сводить с ума, бесить }

{\bf hiding-place } [ˈhaɪ.dɪŋ ˌpleɪs] ~ --- \emph{ укрытие }

{\bf lip } [lɪp] ~ --- \emph{ губа }

{\bf splendour } [ˈsplen.dər] ~ --- \emph{ пышность }

{\bf chaplet } --- \emph{ четки }

{\bf taken } [ˈteɪ.kən] ~ --- \emph{ взятый }

{\bf pearls } --- \emph{ Жемчуг }

{\bf evidently } [ˈev.ɪ.dənt.li] ~ --- \emph{ очевидно }

{\bf averse } [əˈvɜːs] ~ --- \emph{ прочь }

{\bf inclined } [ɪnˈklaɪnd] ~ --- \emph{ наклонный }

{\bf fault } [fɒlt] ~ --- \emph{ вина }

{\bf gossip } [ˈɡɒs.ɪp] ~ --- \emph{ сплетня/сплетничать }

{\bf persuade } [pəˈsweɪd] ~ --- \emph{ убеждать }

{\bf address } [əˈdres] ~ --- \emph{ обращаться/обращение, выступление }

{\bf detached } [dɪˈtætʃt] ~ --- \emph{ отдельный }

{\bf pearl } [pɜːl] ~ --- \emph{ жемчуг, жемчужный }

{\bf intervals } --- \emph{ Интервалы }

{\bf destitute } [ˈdes.tɪ.tʃuːt] ~ --- \emph{ (сильно) нуждающийся }

} \vspace{6mm} {\Large 

 
"My brother and I," said he, "were, as you may imagine, much excited as to the treasure which my father had spoken of. For weeks and for months we dug and delved in every part of the garden without discovering its whereabouts. It was maddening to think that the hiding-place was on his very lips at the moment that he died. We could judge the splendour of the missing riches by the chaplet which he had taken out. Over this chaplet my brother Bartholomew and I had some little discussion. The pearls were evidently of great value, and he was averse to part with them, for, between friends, my brother was himself a little inclined to my father's fault. He thought, too, that if we parted with the chaplet it might give rise to gossip and finally bring us into trouble. It was all that I could do to persuade him to let me find out Miss Morstan's address and send her a detached pearl at fixed intervals so that at least she might never feel destitute."
\\ } 

{\large 

 {\bf companion } [kəmˈpæn.jən] ~ --- \emph{ товарищ, спутник }

{\bf earnestly } --- \emph{ Искренне (не sincerely) }

{\bf extremely } [ɪkˈstriːm.li] ~ --- \emph{ крайне }

{\bf deprecatingly } --- \emph{ неодобрительно }

{\bf trust } [trʌst] ~ --- \emph{ доверять }

{\bf altogether } [ˌɔːl.təˈɡeð.ər] ~ --- \emph{ полностью, всецело }

{\bf plenty } [ˈplen.ti] ~ --- \emph{ множество, избыток }

{\bf ourselves } [ˌaʊəˈselvz] ~ --- \emph{ себя, себе, сами }

{\bf desired } [dɪˈzaɪəd] ~ --- \emph{ Желательный }

{\bf besides } [bɪˈsaɪdz] ~ --- \emph{ кроме }

{\bf taste } [teɪst] ~ --- \emph{ вкус/иметь вкус }

{\bf treated } --- \emph{ лечил (от treat) }

{\bf scurvy } [ˈskɜː.vi] ~ --- \emph{ цинга }

{\bf fashion } [ˈfæʃ.ən] ~ --- \emph{ мода }

{\bf mauvais } --- \emph{ дурным }

{\bf crime } [kraɪm] ~ --- \emph{ преступление }

{\bf French } --- \emph{ французский }

{\bf neat } [niːt] ~ --- \emph{ аккуратный, изящный }

{\bf opinion } [əˈpɪn.jən] ~ --- \emph{ мнение }

{\bf myself } [maɪˈself] ~ --- \emph{ себя, меня самого, себе, сам }

{\bf william } --- \emph{ Уильям }

{\bf extreme } [ɪkˈstriːm] ~ --- \emph{ крайний }

{\bf importance } [ɪmˈpɔː.təns] ~ --- \emph{ важность }

{\bf discovered } --- \emph{ обнаруженный }

{\bf communicated } --- \emph{ Сообщенный }

{\bf remains } [rɪˈmeɪnz] ~ --- \emph{ Остается }

{\bf Norwood } --- \emph{ Норвуд }

{\bf explained } --- \emph{ объяснил (от explain) }

{\bf welcome } [ˈwel.kəm] ~ --- \emph{ желанный, приятный/гостеприимство или приветствие }

} \vspace{6mm} {\Large 

 
"It was a kindly thought," said our companion earnestly; "it was extremely good of you."

The little man waved his hand deprecatingly.

"We were your trustees," he said; "that was the view which I took of it, though Brother Bartholomew could not altogether see it in that light. We had plenty of money ourselves. I desired no more. Besides, it would have been such bad taste to have treated a young lady in so scurvy a fashion. 'Le mauvais go?t m?ne au crime.' The French have a very neat way of putting these things. Our difference of opinion on this subject went so far that I thought it best to set up rooms for myself; so I left Pondicherry Lodge, taking the old khitmutgar and Williams with me. Yesterday, however, I learned that an event of extreme importance has occurred. The treasure has been discovered. I instantly communicated with Miss Morstan, and it only remains for us to drive out to Norwood and demand our share. I explained my views last night to Brother Bartholomew, so we shall be expected, if not welcome, visitors."
\\ } 

{\large 

 {\bf ceased } --- \emph{ Прекратил }

{\bf twitching } --- \emph{ дергающийся }

{\bf luxurious } [lʌɡˈʒʊə.ri.əs] ~ --- \emph{ роскошный }

{\bf remained } --- \emph{ оставшийся }

{\bf mysterious } [mɪˈstɪə.ri.əs] ~ --- \emph{ таинственный, непостижимый }

{\bf remark } [rɪˈmɑːk] ~ --- \emph{ замечание/делать замечание }

{\bf delay } [dɪˈleɪ] ~ --- \emph{ откладывать/отсрочка }

{\bf acquaintance } [əˈkweɪn.təns] ~ --- \emph{ знакомство или знакомый человек }

{\bf deliberately } [dɪˈlɪb.ər.ət.li] ~ --- \emph{ сознательно, обдуманно }

{\bf coil } [kɔɪl] ~ --- \emph{ виток, катушка }

{\bf curtain } [ˈkɜː.tən] ~ --- \emph{ занавеска или занавес }

{\bf topcoat } [ˈtɒp.kəʊt] ~ --- \emph{ пальто }

{\bf astrakhan } [ˈæs.trə.kæn] ~ --- \emph{ каракуль }

{\bf collar } [ˈkɒl.ər] ~ --- \emph{ воротник, ошейник, хомут }

{\bf cuff } [kʌf] ~ --- \emph{ наручники, лёгкий удар }

{\bf button } [ˈbʌt.ən] ~ --- \emph{ кнопка }

{\bf tight } [taɪt] ~ --- \emph{ плотный, тугой/туго }

{\bf spite } [spaɪt] ~ --- \emph{ злоба }

{\bf attire } [əˈtaɪər] ~ --- \emph{ украшение, наряд }

{\bf rabbit-skin } --- \emph{ кролик-кожа }

{\bf hanging } [ˈhæŋ.ɪŋ] ~ --- \emph{ подвешивание }

{\bf lappets } --- \emph{ ушки }

{\bf visible } [ˈvɪz.ə.bəl] ~ --- \emph{ видимый }

{\bf mobile } [ˈməʊ.baɪl] ~ --- \emph{ подвижный }

{\bf peaky } [ˈpiː.ki] ~ --- \emph{ остроконечный }

} \vspace{6mm} {\Large 

 
Mr. Thaddeus Sholto ceased and sat twitching on his luxurious settee. We all remained silent, with our thoughts upon the new development which the mysterious business had taken. Holmes was the first to spring to his feet.

"You have done well, sir, from first to last," said he. "It is possible that we may be able to make you some small return by throwing some light upon that which is still dark to you. But, as Miss Morstan remarked just now, it is late, and we had best put the matter through without delay."

Our new acquaintance very deliberately coiled up the tube of his hookah and produced from behind a curtain a very long befrogged topcoat with astrakhan collar and cuffs. This he buttoned tightly up in spite of the extreme closeness of the night and finished his attire by putting on a rabbit-skin cap with hanging lappets which covered the ears, so that no part of him was visible save his mobile and peaky face.
\\ } 

{\large 

 {\bf somewhat } [ˈsʌm.wɒt] ~ --- \emph{ отчасти, до некоторой степени }

{\bf fragile } [ˈfrædʒ.aɪl] ~ --- \emph{ хрупкий, ломкий }

{\bf compelled } [kəmˈpeld] ~ --- \emph{ Вынужденный }

{\bf valetudinarian } --- \emph{ мнительный }

{\bf await } [əˈweɪt] ~ --- \emph{ ожидать }

{\bf prearranged } [ˌpriː.əˈreɪndʒd] ~ --- \emph{ заранее подготовленный }

{\bf driver } [ˈdraɪ.vər] ~ --- \emph{ водитель, шофер }

{\bf rapid } [ˈræp.ɪd] ~ --- \emph{ быстрый, скорый }

{\bf pace } [peɪs] ~ --- \emph{ шаг/шагать }

{\bf incessantly } --- \emph{ беспрестанно }

{\bf rattle } [ˈræt.əl] ~ --- \emph{ греметь }

} \vspace{6mm} {\Large 

 
"My health is somewhat fragile," he remarked as he led the way down the passage. "I am compelled to be a valetudinarian."

Our cab was awaiting us outside, and our programme was evidently prearranged, for the driver started off at once at a rapid pace. Thaddeus Sholto talked incessantly in a voice which rose high above the rattle of the wheels.
\\ } 

{\large 

 {\bf clever } [ˈklev.ər] ~ --- \emph{ умный }

{\bf fellow } [ˈfel.əʊ] ~ --- \emph{ товарищ или просто чувак }

{\bf treasure } [ˈtreʒ.ər] ~ --- \emph{ сокровище }

{\bf conclusion } [kənˈkluː.ʒən] ~ --- \emph{ вывод }

{\bf somewhere } [ˈsʌm.weər] ~ --- \emph{ где-нибудь, куда-нибудь }

{\bf indoors } [ˌɪnˈdɔːz] ~ --- \emph{ в помещении }

{\bf cubic } [ˈkjuː.bɪk] ~ --- \emph{ кубический }

{\bf measurements } --- \emph{ измерения }

{\bf everywhere } [ˈev.ri.weər] ~ --- \emph{ везде }

{\bf unaccounted } [ˌʌnəˈkaʊntɪd] ~ --- \emph{ неучтенный }

{\bf height } [haɪt] ~ --- \emph{ высота }

{\bf seventy-four } --- \emph{ семьдесят четыре }

{\bf height } [haɪt] ~ --- \emph{ высота }

{\bf ascertained } --- \emph{ Установленный }

{\bf boring } [ˈbɔː.rɪŋ] ~ --- \emph{ скучный }

{\bf seventy } [ˈsev.ən.ti] ~ --- \emph{ семьдесят }

{\bf knocked } --- \emph{ стучать (2 форма) *n- }

{\bf therefore } [ˈðeə.fɔːr] ~ --- \emph{ поэтому, следовательно }

{\bf lath } [lɑːθ] ~ --- \emph{ планка }

{\bf plaster } [ˈplɑː.stər] ~ --- \emph{ штукатурка/штукатурить, оклеивать, покрывать }

{\bf highest } --- \emph{ наибольший }

{\bf garret } [ˈɡær.ət] ~ --- \emph{ мансарда, чердак }

{\bf seal } [siːl] ~ --- \emph{ печать/опечатывать }

{\bf known } [nəʊn] ~ --- \emph{ известный }

{\bf treasure-chest } --- \emph{ сундук с сокровищами }

{\bf raft } [rɑːft] ~ --- \emph{ плот }

{\bf computes } --- \emph{ вычисляет }

{\bf jewels } --- \emph{ драгоценности }

{\bf sterling } [ˈstɜː.lɪŋ] ~ --- \emph{ стерлинг }

} \vspace{6mm} {\Large 

 
"Bartholomew is a clever fellow," said he. "How do you think he found out where the treasure was? He had come to the conclusion that it was somewhere indoors, so he worked out all the cubic space of the house and made measurements everywhere so that not one inch should be unaccounted for. Among other things, he found that the height of the building was seventy-four feet, but on adding together the heights of all the separate rooms and making every allowance for the space between, which he ascertained by borings, he could not bring the total to more than seventy feet. There were four feet unaccounted for. These could only be at the top of the building. He knocked a hole, therefore, in the lath and plaster ceiling of the highest room, and there, sure enough, he came upon another little garret above it, which had been sealed up and was known to no one. In the centre stood the treasure-chest resting upon two rafters. He lowered it through the hole, and there it lies. He computes the value of the jewels at not less than half a million sterling."
\\ } 

{\large 

 {\bf mention } [ˈmen.ʃən] ~ --- \emph{ упоминать/упоминание }

{\bf gigantic } [ˌdʒaɪˈɡæn.tɪk] ~ --- \emph{ гигантский }

{\bf open-eyed } [ˌəʊ.pənˈaɪd] ~ --- \emph{ с широко раскрытыми глазами }

{\bf secure } [sɪˈkjʊər] ~ --- \emph{ безопасный, спокойный/обезопасить, сохранить }

{\bf needy } [ˈniː.di] ~ --- \emph{ нуждающийся }

{\bf governess } [ˈɡʌv.ən.əs] ~ --- \emph{ гувернантка }

{\bf richest } --- \emph{ богатейший }

{\bf heiress } [ˈeə.res] ~ --- \emph{ наследница }

{\bf loyal } [ˈlɔɪ.əl] ~ --- \emph{ преданный }

{\bf rejoice } [rɪˈdʒɔɪs] ~ --- \emph{ радовать(ся) }

{\bf ashamed } [əˈʃeɪmd] ~ --- \emph{ пристыженный, стыдящийся }

{\bf soul } [səʊl] ~ --- \emph{ душа }

{\bf stammer } [ˈstæm.ər] ~ --- \emph{ заикаться, запинаться }

{\bf halt } [hɒlt] ~ --- \emph{ останавливаться, делать привал/остановка, привал }

{\bf congratulation } [kənˌɡrætʃ.əˈleɪ.ʃən] ~ --- \emph{ поздравление }

{\bf downcast } [ˈdaʊn.kɑːst] ~ --- \emph{ удрученный }

{\bf droop } [druːp] ~ --- \emph{ поникать }

{\bf deaf } [def] ~ --- \emph{ глухой }

{\bf babble } [ˈbæb.əl] ~ --- \emph{ бормотание, болтовня }

{\bf confirmed } [kənˈfɜːmd] ~ --- \emph{ закоренелый, убеждённый }

{\bf hypochondriac } --- \emph{ ипохондрик }

{\bf dreamily } [ˈdriː.məl.i] ~ --- \emph{ мечтательно }

{\bf conscious } [ˈkɒn.ʃəs] ~ --- \emph{ сознательный, сознающий }

{\bf pour } [pɔːr] ~ --- \emph{ лить }

{\bf interminable } [ɪnˈtɜː.mɪ.nə.bəl] ~ --- \emph{ бесконечный }

{\bf symptoms } --- \emph{ симптомы }

{\bf imploring } --- \emph{ умоляющий }

{\bf composition } [ˌkɒm.pəˈzɪʃ.ən] ~ --- \emph{ сочинение, композиция }

{\bf innumerable } [ɪˈnjuː.mər.ə.bəl] ~ --- \emph{ бесчисленный }

{\bf quack } [kwæk] ~ --- \emph{ крякать }

{\bf nostrums } --- \emph{ панацея }

{\bf bore } [bɔːr] ~ --- \emph{ зануда }

{\bf leather } [ˈleð.ər] ~ --- \emph{ кожа (материал) }

{\bf pocket } [ˈpɒk.ɪt] ~ --- \emph{ карман/класть в карман }

{\bf trust } [trʌst] ~ --- \emph{ доверять }

{\bf declares } --- \emph{ объявляет }

{\bf overheard } --- \emph{ подслушал }

{\bf caution } [ˈkɔː.ʃən] ~ --- \emph{ осторожность, предусмотрительность }

{\bf castor-oil } [ˌkɑː.stər ˈɔɪl] ~ --- \emph{ касторка }

{\bf recommended } [ˌrekəˈmendɪd] ~ --- \emph{ Рекомендуемый }

{\bf strychnine } [ˈstrɪk.niːn] ~ --- \emph{ стрихнин }

{\bf dos } [duːz] ~ --- \emph{ ДОС }

{\bf sedative } [ˈsed.ə.tɪv] ~ --- \emph{ успокоительное }

{\bf relieved } [rɪˈliːvd] ~ --- \emph{ избавленный }

{\bf jerk } [dʒɜːk] ~ --- \emph{ резкий толчок, подергивание; резко толкать, дергать }

{\bf coachman } --- \emph{ кучер }

{\bf sprang } [spræŋ] ~ --- \emph{ Прыгал (от слова spring) }

} \vspace{6mm} {\Large 

 
At the mention of this gigantic sum we all stared at one another open-eyed. Miss Morstan, could we secure her rights, would change from a needy governess to the richest heiress in England. Surely it was the place of a loyal friend to rejoice at such news, yet I am ashamed to say that selfishness took me by the soul and that my heart turned as heavy as lead within me. I stammered out some few halting words of congratulation and then sat downcast, with my head drooped, deaf to the babble of our new acquaintance. He was clearly a confirmed hypochondriac, and I was dreamily conscious that he was pouring forth interminable trains of symptoms, and imploring information as to the composition and action of innumerable quack nostrums, some of which he bore about in a leather case in his pocket. I trust that he may not remember any of the answers which I gave him that night. Holmes declares that he overheard me caution him against the great danger of taking more than two drops of castor-oil, while I recommended strychnine in large doses as a sedative. However that may be, I was certainly relieved when our cab pulled up with a jerk and the coachman sprang down to open the door.
\\ } 


\vspace{15mm}

\section{Информация для читателя}
\vspace{4mm}

Основной метод изучения иностранных слов --- это их активное использование в контексте.

Перед отрывками текста дается список слов, не входящих в список 1000 самых используемых слов английского языка. Соответственно, для прочтения представленных здесь книг достаточно запомнить эти самые 1000 слов.

\vspace{4mm}
{ \large
Список 1000 самых употребимых слов английского, а также другие книги можно скачать на нашем веб-сайте \url{http://english.proger.site/}
}


\end{document}
